# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r read, echo=T}

activdata<-read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))
head(activdata)
```


## What is mean total number of steps taken per day?
```{r meand, echo=T, fig.height=4}
library(data.table)
activtable<-data.table(activdata)
plotdata<-na.omit(activtable[,sum(steps), by = "date"])
hist(plotdata$V1)
meandata<-mean(plotdata$V1, na.rm=T)
meddata<-median(plotdata$V1, na.rm=T)
print(meandata)
print(meddata)

```

## What is the average daily activity pattern?
```{r avgd, echo=T, fig.height=4}
cleandata<-na.omit(activtable)
avgdata<-cleandata[,mean(steps), by = "interval"]
plot(avgdata$interval,avgdata$V1, type ="l")
m<-max(avgdata$V1)
maximumd<-subset(avgdata, avgdata$V1 == m) 
print(maximumd)

```
## Imputing missing values

```{r miss, echo=T, warning =FALSE}

activdata<-read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))
nasum<-sum(is.na(activdata))
print(nasum)


activdata<-read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))
library(zoo)
fulltable<-activdata
fulltable[1:10,1]<-0
z<-zoo(fulltable[,1], as.Date(fulltable$date), fulltable$interval)

w<-weekdays(fulltable$date)


fulltable<-na.aggregate(z, format, "%d", FUN = mean, na.rm=T)
fulltab<-as.data.frame(fulltable)
activtab<-activdata[,2:3]
finaltab<-cbind(fulltab, activtab)
colnames(finaltab)<-c("steps", "date", "interval")


library(data.table)
activq<-data.table(finaltab)
plotq<-na.omit(activq[,sum(steps), by = "date"])
hist(plotq$V1)
meanq<-mean(plotq$V1, na.rm=T)
medq<-median(plotq$V1, na.rm=T)
print(meanq)
print(medq)

```
### The value after repalcing NA are lower than earlier as some zeros are included in overall mean to lower the mean and median

## Are there differences in activity patterns between weekdays and weekends?

```{r week, echo=T}

p<-nrow(finaltab)


for (i in 1:p){
  
if(weekdays(finaltab[i,2]) == "Saturday"){
  
finaltab[i,4]<-"weekend"
}
else{
  if(weekdays(finaltab[i,2]) == "Sunday"){
    finaltab[i,4]<-"weekend"
    }
  else{
finaltab[i,4]<-"weekday"
}
}
}

library(data.table)
library(ggplot2)
library(reshape2)
library(gridExtra)
finalq<-data.table(finaltab)
plotq<-finalq[,mean(steps), by = "interval,V4"]
f<-ggplot(plotq) + facet_grid(~V4, as.table =T) + geom_line(aes(x=interval, y=V1, group = V4))
grid.arrange(f, nrow=2)

```
